A Framework for Predictable Actor-Critic Control

### 2022-05-09 documentation/paper.tex
minor changes

        To add planning and commitment to an agent, the proposed framework requires the handling three basic tasks. The first is choosing an action (\(\pi(s)\)), the second task is determining the long term value of an action in a given state \(Q(s,a)\), and the third is to create a projection of planned actions given the current state and action. The first two tasks can easily be addressed with an actor and critic from a standard SAC design. For the state prediction task, a simple dynamics model is used for predicting state values given a current state and action. To generate an \textit{imaginary plan}~\cite{racaniere2017imagination}, the dynamics model is used to produce an estimated next state, which is then fed to actor to produce an action. This process is repeated, to generate plans of arbitrary length. These are specific choices, and the approach should hold for any agent capable of the three core tasks.

@@

    \section{Related Work}

        This paper proposes extending the common actor-critic framework to compute and follow $n$-step plans. In order to do so, we suggest training a model approximator for the underlying state transition probabilities ($P$).   
        Training a model of the environment, i.e., approximating $P(s'|s,a)$ based on observed transitions, is commonly done as part of a \textit{model-based RL}~\cite{kaiser2019model,xu2020prediction} approach. Such a model can be used to produce imaginary trajectories~\cite{racaniere2017imagination} that improve training sample efficiency in many domains. The imagined trajectory is a simulated future trajectory based on the learned model and current policy. We differentiate between the imaginary trajectory, which is a set of predicted $(s,a,r,s')$ transitions, and the imaginary plan, which is the affiliated sequence of actions in the imaginary trajectory.

        For computing the imaginary trajectory, Racaniere et al.~\citeyear{racaniere2017imagination} trained an environment model with a recurrent architecture. The proposed environment model extended on the notion of action-conditional next-step predictors~\cite{oh2015action,chiappa2017recurrent,leibfried2017deep}, which predicts the next state, and potentially the reward, received after following a given action at a given state (or history of states).
        It is important to note that modeling errors can compound over long trajectories resulting in an unlikely imaginary trajectory and an affiliated risky imaginary plan. Such modeling errors were shown to be common in complex domains~\cite{talvitie2014model,talvitie2015agnostic}. \JA{present as baseline method}

@@

        %The agent generates an imaginary $n$-step plan of its next n actions it intends to take (including the action it plans to take in the current timestep). In this work, when we say plan, we are referring to such an imaginary plan. 
        Kim et al.~\citeyear{kim2020communication} proposed to utilize imaginary plans towards intention sharing in multiagent RL scenarios. It was demonstrated that broadcasting agents' future intentions (the imagined plans) improve coordination and overall performance over sharing just the current and past states of agents~\cite{foerster2016learning,sukhbaatar2016learning,jiang2018learning,das2019tarmac}.

@@

        Perpetually changing the predicted (imaginary) trajectory, however, does not lend well to predicting which actions an agent will take in the future.
        In our approach, Predictable Actor-Critic (PPAC), we suggest a method of merging these two goals in a generalized framework by reviewing the current plan at each new state and preserving planned steps for as long as some performance degradation threshold is met.
        A predictable control plan is therefore a trade-off between maximizing the length of predicted plans and minimizing performance degradation. We utilize values learned by the critic component of a $Q$-learning actor-critic RL algorithm to evaluate trajectories imagined through an approximated model. \JA{introduce coach}

>>>>

        To add planning and commitment to an agent, the proposed framework requires the handling three basic tasks. The first is choosing an action (\(\pi(a | s)\)), the second task is determining the long term value of an action in a given state \(Q(s,a)\), and the third is to create a projection of planned actions given the current state and action \(P(s' | s,a)\). The first two tasks can easily be addressed with an actor and critic from a standard SAC design. For the state prediction task, a simple dynamics model is used for predicting state values given a current state and action. To generate an \textit{imaginary plan}~\cite{racaniere2017imagination}, the dynamics model is used to produce an estimated next state, which is then fed to actor to produce an action. This process is repeated, to generate plans of arbitrary length. These are specific choices, and the approach should hold for any agent capable of the three core tasks.

@@

    \section{Related work}

        Approximating $P(s'|s,a)$ based on observed transitions, is commonly done as part of a \textit{model-based RL}~\cite{kaiser2019model,xu2020prediction} approach. Such a model can be used to produce imaginary trajectories~\cite{racaniere2017imagination} that improve training sample efficiency in many domains. The imagined trajectory is a simulated future trajectory based on the learned model and current policy. We differentiate between the imaginary trajectory, which is a set of predicted $(s,a,r,s')$ transitions, and the imaginary plan, which is simply a sequence of actions.

        For computing the imaginary trajectory, Racaniere et al.~\citeyear{racaniere2017imagination} trained an environment model with a recurrent architecture. The proposed environment model extended on the notion of action-conditional next-step predictors~\cite{oh2015action,chiappa2017recurrent,leibfried2017deep}, which predicts the next state, and potentially the reward, received after following a given action at a given state (or history of states).
        It is important to note that modeling errors can compound over long trajectories resulting in an unlikely imaginary trajectory and a risky imaginary plan. Such modeling errors were shown to be common in complex domains~\cite{talvitie2014model,talvitie2015agnostic}.

@@

        %The agent generates an imaginary $n$-step plan of its next n actions it intends to take (including the action it plans to take in the current timestep). In this work, when we say plan, we are referring to such an imaginary plan. 
        Kim et al.~\citeyear{kim2020communication} proposed to utilize imaginary plans towards intention sharing in multiagent RL scenarios. It was demonstrated that broadcasting agents' imagined plans improved coordination and overall performance when compared to broadcasting the current and past states of agents~\cite{foerster2016learning,sukhbaatar2016learning,jiang2018learning,das2019tarmac}.

@@

        Perpetually changing the predicted (imaginary) trajectory, however, does not lend well to predicting which actions an agent will take in the future.
        In our approach, Predictable Actor-Critic (PPAC), we suggest a method of merging these two goals in a generalized framework by reviewing the current plan at each new state and preserving planned steps for as long as some performance degradation threshold is met.
        A predictable control plan is therefore a trade-off between maximizing the length of predicted plans and minimizing performance degradation. We utilize values learned by the critic component of a $Q$-learning actor-critic RL algorithm to evaluate trajectories imagined through an approximated model.
